NMRlipids IV: Headgroup \& glycerol backbone structures, and cation binding in bilayers with PS lipids

### 2018-09-17 Manuscript/manuscriptPS.tex
Update discussion about PS headgroup structure based on new figures by P. Buslaev

The headgroup and glycerol backbone of PS lipids show wide variety between different simulation
models in the order parameters and structures (Figs. \ref{HGorderParametersPS} and \ref{HGstructuresPS}),
as previously observed also for PC lipids. The models for PS lipids perform generally less well than for
PC lipids in the previous study
(Figs. 2 and 4 in Ref.~\cite{botan15} vs. Figs.~\ref{HGorderParametersPS} and \ref{comparisonTablePS}).
Therefore, interpretation of structural differences between PC and PS headgroups
from simulations is not straightforward.

The best performing models,
Slipids, CHARMM36 and CHARMM36ua, reproduce the larger forking of the $\alpha$-carbon 
and the Slipids model reproduces also the lower of the $\beta$-carbon order parameter
when comparing the PS results to PC (Fig. 2 in Ref.~\citenum{botan15} vs. Fig. \ref{HGorderParametersPS}).
Interestingly, the headgroup structures of these models share significant similarities
and are distinct from other models (Figs.~\ref{HGstructuresPS} and \ref{dihedralsHG}). 
The distribution of the dihedral angle for C$_\alpha$-C$_\beta$-C$_\gamma$-O$_\gamma$ shows
a single narrow maximum close to 120$^{\circ}$ in the best three models, while other models give several maxima
in different locations (Fig. \ref{dihedralsHG}). The restricted motion is also
visible in the sampled conformations (Fig.~\ref{HGstructuresPS}) and might
manifest the increased rigidity anticipated from the early experimental studies \cite{browning80,buldt81}.
Another distinct feature in the three best performing models is the distribution
of N-C$_\beta$-C$_\alpha$-O$_\alpha$ dihedral, which gives a large maxima close to 300$^{\circ}$
and a smaller maxima close to 60$^{\circ}$. This is also visible in the sampled
structures (Fig.~\ref{HGstructuresPS}).
\todo{There seems to be discrepancy between structures and dihedral distributions for CHARMM36ua}.
The glycerol backbone order parameters of C$_2$ and C$_3$ from Slipids simulations
differ significantly from the other simulation results and experiments (Fig. \ref{HGorderParametersPS}),
as observed previously also for PC lipids \cite{botan15}.
The origin of this difference is more difficult to track without more elaborate analysis,
because different models show very complicated patterns of distinct structures
in the glycerol backbone region (Figs.~\ref{HGstructuresPS} and \ref{dihedralsHG}).

>>>>

The headgroup and glycerol backbone of PS lipids show wide variety between different simulation
models in the order parameters and structures (Figs. \ref{HGorderParametersPS} and \ref{HGstructuresPS}),
as previously observed also for PC lipids \cite{botan15}. The models for PS lipids perform generally less well than for
PC lipids in the previous study
(Figs. 2 and 4 in Ref.~\cite{botan15} vs. Figs.~\ref{HGorderParametersPS} and \ref{comparisonTablePS}).
Therefore, interpretation of structural differences between PC and PS headgroups
from simulations is not straightforward.

The best performing models,
Slipids, CHARMM36 and CHARMM36ua, reproduce the larger forking of the $\alpha$-carbon 
and the Slipids model reproduces also the lower of the $\beta$-carbon order parameter
when comparing the PS results to PC (Fig. 2 in Ref.~\citenum{botan15} vs. Fig. \ref{HGorderParametersPS}).
Interestingly, the dihedral angle distributions of C$_\alpha$-C$_\beta$-C$_\gamma$-O$_\gamma$ show
a single narrow maximum close to 120$^{\circ}$ in the best three models, while other models give several maxima
in different locations (Fig. \ref{dihedralsHG}). The restricted motion is also
visible in the sampled conformations (Fig.~\ref{HGstructuresPS}) and might
manifest the increased rigidity anticipated from the early experimental studies \cite{browning80,buldt81}.
\todo{We should consider making the conformation figures from CHARMM36 and Slipids POPC, and compare these
  to the results from POPS.}

\todo{Maybe we should move the glycerol backbone results to SI and focus here on the headgroup,
and especially to the difference between PC and PS.}
The glycerol backbone order parameters of C$_2$ and C$_3$ from Slipids simulations
differ significantly from the other simulation results and experiments (Fig. \ref{HGorderParametersPS}),
as observed previously also for PC lipids \cite{botan15}.
The origin of this difference is more difficult to track without more elaborate analysis,
because different models show very complicated patterns of distinct structures
in the glycerol backbone region (Figs.~\ref{HGstructuresPS} and \ref{dihedralsHG}).
